Towards Detecting Stalkerware Through Network Traces

### 2019-03-18 paper/main.tex
edit traffic result

However, some characteristics of the stalkerware apps themselves may aid us in 
detecting them. For example, similar to malware, some groups of apps appear to 
be repackaged versions of each other. For example, PhoneSpector, 
HighsterMobile, AutoForward, EasySpy and Surepoint Spy all have separate 
websites, but the appearance of each of these sites is nearly identical. Upon 
closer inspection, although multiple companies are listed as the makers of the 
apps (Powerline Group Corp., ILF Mobile Apps, and PhoneSpector LLC), all have 
the same physical address listed. At least one blog that is supposedly run by 
an independent part reviewed apps that coincidentally were made by the same 
company\cite{noauthor_best_nodate}. Other app sites appear to be similar to each other in other ways, 
with many having a large panel with a darkened image background extolling the 
features of the app above two buttons, one for a "live demo" and the other to 
"buy now." This is somewhat circumstantial evidence - it is entirely possible 
that these manufacturers simply chose similar web templates, or copied each 
other's site in an attempt to undercut competition. However, if multiple groups 
of stalkerware apps are in fact made by only a few manufacturers, it may 
facilitate app identification.

After our investigation on these apps, we can confirm that, at least from the few that we have selected, many stalkerware apps are related in some fashion to each other. The most notable being the group of apps stated above, and a single app called spy2mobile. Spy2mobile is an interesting case because it fits the description of a generic stalkerware app so closely that it is the perfect example of spyware on the Play Store. Moreover, this single app has been repackaged multiple times. We have discovered that there are at least 6 copies of the app on the Play store, all with different developers and different names. While this is still circumstantial evidence that most stalkerware often contains copies of each other, we can at least conclue that our forty-five apps constitute a larger percentage than anticipated.

@@

What causes most stalkerware to send packets? Our captures and testing show that there are three main categories of updating tracker state: manual, timed, and on tracked feature. Most free apps fall into the manual update method, where there is a button on a website or on a phone app that will send a signal packet to the stalkerware app that initalizes the upload. Most paid apps also include a timed auto-update feature, where an update will be sent every few minutes. A very small subset of both advertised that it receives a "real time notification" every time email, texts, calls, or GPS, was changed or added to. Our results show that none of the stalkerware apps were able to send in real time.

@@

\subsection{Network Signatures}

@@

The majority of stalkerware apps communicate though HTTP or HTTPS.

>>>>

However, some characteristics of the stalkerware apps themselves may aid us in 
detecting them. For example, similar to malware, some groups of apps appear to 
be repackaged versions of each other. For example, PhoneSpector, 
HighsterMobile, AutoForward, EasySpy and Surepoint Spy all have separate 
websites, but the appearance of each of these sites is nearly identical. Upon 
closer inspection, although multiple companies are listed as the makers of the 
apps (Powerline Group Corp., ILF Mobile Apps, and PhoneSpector LLC), all have 
the same physical address listed. At least one blog that is supposedly run by 
an independent part reviewed apps that coincidentally were made by the same 
company [9]. Other app sites appear to be similar to each other in other ways, 
with many having a large panel with a darkened image background extolling the 
features of the app above two buttons, one for a “live demo” and the other to 
“buy now.” This is somewhat circumstantial evidence - it is entirely possible 
that these manufacturers simply chose similar web templates, or copy each 
other’s site in an attempt to undercut competition. However, if multiple groups 
of stalkerware apps are in fact made by only a few manufacturers, it may 
facilitate app identification. Similarities in network traces may indicate that 
the code used in the apps was made by the same developers.

@@

@@

@@

The majority of stalkerware apps communicate though HTTP or HTTPS, while some use their own protocols. We examined the domain, SNI, User-Agent, and HTTP host, for each app. The words “spy,” “find,” and “track” appeared frequently in various fields of the traffic. The most common place to find them was in the destination domain, which we found using Wireshark’s reverse DNS lookup tool. 12 of the 26 apps that we eventually found signatures for contained these spyware keywords in their domains. We identified the following generalized signatures that can be used to identify most of our apps given nothing but information available on the Google Play store.
